Cartogram Mapping and its Application to Cancer Data Visualisation

### 2019-08-01 review_article.Rmd
Examples of cartograms in literature

<!-- psychology -->
Utilising the familiar state boundaries can make a map intuitive to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data, i.e. how cancer rate differs across the states.
The familiarity of the geography is a worthy consideration when presenting results of spatial analysis.
While the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time. 
The different population and geographical sizes of administrative areas can attract attention to the shades of the unpopulated but large areas [@EI]. 
Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].

@@

<!-- Displaying of cancer data on choropleths -->

In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
Choropleth maps can be useful devices for communicating information to public on a familiar map base.
A cancer atlas is a choropleth map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries. @MACM provide the definition and reports that cancer atlases began with Haviland's maps in 1875.
The data collection methods of cancer mortality rates across regions, and the administrative control within regions lends itself to choropleth visualisation. 
The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
Early work in US cancer atlases can be attributed to Burbank ???, and UK cancer atlases to Howe 1963 ???.
The choropleth maps presented levels via hatchings or dots on a black and white scale.
These atlases were key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets @MACM.

<!-- Measures reported in cancer atlases -->

Almost 100 years of cancer mapping in the United States and the United Kingdom has seen increased effectiveness in the presentation cancer statistics.
Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the the higher prevalence of cancers in older populations.
Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s. Table 1 presents summarises the measures presented in published cancer atlases, and provides a definition of each measure.

@@

An Online Census Atlas for everyone [@AOCAFE] uses familiar choropleth, produced using generalized boundaries for six higher level geographies. Great Britain is recognizable and it is relatively easy to identify the higher geography levels such as Counties and Local Authorities.

Other characteristics of the population
may also serve as the base for adjustment; for
example, sex, race, parity, socio-economic level,
or immune status.

@@

>>>>

<!-- psychology -->
Utilising the familiar state boundaries can make a map intuitive to read [@CIBMUK], and allow viewers to visually infer the spatial relationships in the data, i.e. how cancer rate differs across the states.
The familiarity of the geography is a worthy consideration when presenting results of spatial analysis. Just as geographers are no longer the only creators of maps, @CPISACA suggests the audiences of spatial health data analysis have extended beyond researchers to the public, policymakers and the media.
While the areas are recognisable shapes, they are often politically driven boundaries with individual areas being of non-uniform size, containing different population densities and subject to change over time. 
The different population and geographical sizes of administrative areas can attract attention to the shades of the unpopulated but large areas [@EI]. 
Choropleths can inhibit visual inference when presenting human related statistics as the display may draw attention from the 'potentially more important results in the more populous communities' [@SE].

@@

<!-- Displaying of cancer data on choropleths -->
Choropleth maps can be useful devices for communicating information to public on a familiar map base.
In epidemiology, choropleths are often used as a tool to study the spatial distribution of cancer incidence and mortality. 
A cancer atlas is a choropleth map, or collection of maps, representing cancer incidence and mortality for a country, or group of countries. @MACM provide the definition and reports that cancer atlases began with Haviland's maps in 1875.
The data collection methods of cancer mortality rates across regions, and the administrative control within regions lends itself to choropleth visualisation. 
The increasing development and use of disease maps can be attributed to the availability of geographic information system software [@SE].
Early work in US cancer atlases can be attributed to Burbank ???, and UK cancer atlases to Howe 1963 ???.
The choropleth maps presented levels via hatchings or dots on a black and white scale.
These atlases were key to developing hypotheses regarding areas with unusually high rates, geographic correlations, work related exposures, and high risk diets @MACM.

<!-- Measures reported in cancer atlases -->
Almost 100 years of cancer mapping in the United States and the United Kingdom has seen increased effectiveness in the presentation cancer statistics.
Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the the higher prevalence of cancers in older populations.
Howe [-@HEDP] describes Stock's development of the standardised mortality ratios through the 1930s. Table 1 presents summarises the measures presented in published cancer atlases, and provides a definition of each measure.

@@

<!-- Cancer atlas literature -->

@@

Other characteristics of the population
may also serve as the base for adjustment; for
example, sex, race, parity, socio-economic level,
or immune status.
-->
